Fast Hyper-Parameter Selection for Graph Kernels via Subsampling and Multiple Kernel Learning

### 2016-11-25 paper.tex
Typos

%\section{Kernel for structures and Multiple Kernel Learning}
\section{Background}
A kernel method is composed by the following two modules.
A learning algorithm, that expresses the solution only via dot products between training examples.
A kernel function ${k}: \mathcal{X} \times \mathcal{X} \rightarrow \mathbb{R}$ that is a symmetric positive semi-definite function that corresponds to a dot product in a Reproducing Kernel Hilbert Space (RKHS), i.e. there exists a $\phi: \mathcal{X} \rightarrow \mathcal{K} \subseteq \mathbb{R}^D$, where $\mathcal{X}$ is the input space and $\mathcal{K}$ is an Hilbert space (commonly referred to as \textit{feature space}), such that $k(x_i,x_j) = \langle \phi(x_i),\phi(x_j) \rangle$ with $x_i,x_j \in \mathcal{X}$. %~\cite{559923}.
%\textcolor{red}{TODO definire kernel matrix}
A kernel (or \textit{Gram}) matrix for a set of examples $Tr=\{x_1, \ldots, x_l\}$ is an $l \times l$ matrix $G$ with entries $\{G\}_{ij}=k(x_i,x_j)$.
When dealing with structured data, the choice of the kernel to adopt is a key step. Depending on the type of data under consideration, different proposals are available in literature, each one with a different trade-off between predictive performance and computational time (of course, this trade-off depends from the considered task).

In this paper we focus on graphs, i.e. we consider each example to be a different graph.
A graph is a tuple $G=(V_G,E_G,L_G)$, where $V_G=\{v_1,\ldots,v_n\}$ is the set of vertices, $E_G=\{(u,v) | u,v \in V_G\}$ is the set of edges (with $|E_G|=m$), and $L_G: V_G \rightarrow \Sigma$ is a function mapping each vertex to a discrete label in a fixed alphabet $\Sigma$.
%A graph is undirected if $(u,v)\in E_g \Rightarrow (v,u) \in E_G$, otherwise it is directed.
In this setting, there are several graph kernels defined in literature based on random walks~\cite{Kashima03marginalizedkernels,Mahe2004} ($O(n^3)$ computational complexity), shortest paths~\cite{Kriegel05shortestpath} ($O(n^4)$ computational complexity), subgraphs up to a fixed size $h$~\cite{Shervashidze2009} ($O(n^h)$ complexity). The problem with these kernels is the high computational complexity, that makes them inapplicable to several real-world datasets.
Recently, different almost linear time graph kernels have been defined in literature \cite{Heinonen2009,Costa2010,Shervashidze2011,Dasan2012,DaSanMartino2016}, and these are the ones we focus on in this paper.
In the following we will give some details about the considered alternatives.
Each kernel restricts to specific kinds of subgraphs, up to a certain dimension (that, in general, is a kernel parameter).
Note that it is out of the scope of this paper to give an extensive comparison among the various graph kernels.

@@

In~\cite{Costa2010} the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK), an extension of the WL Fast Subtree Kernel, is presented. This kernel computes the exact matches between pairs of subgraphs with controlled size and distance (hyper-parameters $r$ and $d$, respectively). 
The kernel has $O(|V| |V_h| |E_h| \log |E_h|)$ time complexity, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
% The complexity of the kernel is $O(|V| |V_h| |E_h| \log |E_h|)$, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
The authors state that, for small values of the subgraph size and distance, the complexity of the kernel becomes in practice linear.

>>>>

%\section{Kernel for structures and Multiple Kernel Learning}
\section{Background}
A kernel method is composed by the following two modules.
A learning algorithm, that expresses the solution only via dot products between training examples.
A kernel function ${k}: \mathcal{X} \times \mathcal{X} \rightarrow \mathbb{R}$ that is a symmetric positive semi-definite function that corresponds to a dot product in a Reproducing Kernel Hilbert Space (RKHS), i.e. there exists a $\phi: \mathcal{X} \rightarrow \mathcal{K} \subseteq \mathbb{R}^D$, where $\mathcal{X}$ is the input space and $\mathcal{K}$ is an Hilbert space (commonly referred to as \textit{feature space}), such that $k(x_i,x_j) = \langle \phi(x_i),\phi(x_j) \rangle$ with $x_i,x_j \in \mathcal{X}$. %~\cite{559923}.
%\textcolor{red}{TODO definire kernel matrix}
A kernel (or \textit{Gram}) matrix for a set of examples $Tr=\{x_1, \ldots, x_l\}$ is an $l \times l$ matrix $G$ with entries $\{G\}_{ij}=k(x_i,x_j)$.
When dealing with structured data, the choice of the kernel to adopt is a key step. Depending on the type of data under consideration, different proposals are available in literature, each one with a different trade-off between expressiveness and efficiency.
%(of course, this trade-off depends from the considered task).

In this paper we focus on graphs, i.e. we consider each example to be a different graph.
A graph is a tuple $G=(V_G,E_G,L_G)$, where $V_G=\{v_1,\ldots,v_n\}$ is the set of vertices, $E_G=\{(u,v) | u,v \in V_G\}$ is the set of edges (with $|E_G|=m$), and $L_G: V_G \rightarrow \Sigma$ is a function mapping each vertex to a discrete label in a fixed alphabet $\Sigma$.
%A graph is undirected if $(u,v)\in E_g \Rightarrow (v,u) \in E_G$, otherwise it is directed.
In this setting, there are several graph kernels defined in literature based on random walks~\cite{Kashima03marginalizedkernels,Mahe2004} ($O(n^3)$ computational complexity), shortest paths~\cite{Kriegel05shortestpath} ($O(n^4)$ computational complexity), subgraphs up to a fixed size $h$~\cite{Shervashidze2009} ($O(n^h)$ complexity). The problem with these kernels is the high computational complexity, that makes them inapplicable to several real-world datasets.
Recently, different almost linear time graph kernels have been defined in literature \cite{Heinonen2009,Costa2010,Shervashidze2011,Dasan2012,DaSanMartino2016}, and these are the ones we focus on in this paper.
%In the following we will give some details about the considered alternatives.
Each one of these kernels considers only specific kinds of subgraphs, up to a certain dimension (that, in general, is a kernel parameter).
Note that it is out of the scope of this paper to give an extensive comparison among the various graph kernels.

@@

In~\cite{Costa2010} the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) is presented. This kernel computes the exact matches between pairs of subgraphs with controlled size and distance (hyper-parameters $r$ and $d$, respectively). 
The kernel has $O(|V| |V_h| |E_h| \log |E_h|)$ time complexity, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
% The complexity of the kernel is $O(|V| |V_h| |E_h| \log |E_h|)$, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
The authors state that, for small values of the subgraph size and distance, the complexity of the kernel becomes in practice linear.
